---
title: "Using OLS Regression to Predict Median House Values in Philadelphia" 
author: names 
date: date 
output: 
  html_document: 
    toc: true 
    toc_float: true 
    toc_depth: 6
    code_folding: "hide"
    theme: sandstone
    highlight: textmate
editor_options: 
  markdown: 
    wrap: 72
---

------------------------------------------------------------------------

------------------------------------------------------------------------

```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set()

#--- load libraries ----


library(tidyr)
library(dplyr)
library(DAAG)
library(car)  #to calculate VIF
library(MASS)
library(rsq)
library(kableExtra)
library(tidyverse) #for ggplot
library(sf) #for maps
library(cowplot) #for plotgrid
library(classInt)#for jenks breaks
library(rgdal)
library(ggplot2)
library(RColorBrewer)
#install.packages("broom") *NEW*
library(broom)
library(stargazer)



options(scipen=999)

#---- Step 1: Upload Data ----

ourdata <- read.csv("https://raw.githubusercontent.com/bri-ne/MUSA500_Assignment_1/main/RegressionData.csv")

```

#### **Introduction**

*intro paragraphs (2)*

#### **Methods**

###### **a. Data Cleaning**

Our study relies on census data, that includes observations for various
demographic variables at the census block group level. We began with a
data set of 1,816 observation and filtered out block groups with small
populations (\<40), block groups with no housing units, and block groups
with a median house value less than \$10,000. Additionally, a two
separate block groups in North Philadelphia were removed due to a very
high median house value(greater than \$80,000) and very low median
household income (less than \$8,000). We were left with 1,720
observation to use in our analysis.

###### **b. Exploratory Analysis**

The first step we took in analyzing our data set, was to calculate
summary statistics. These statistics include the mean and standard
deviation of both our dependent variable Median House Value (MEDHVAL)
and four independent predictors:

-   Number of Households Living Below the Poverty Line (NBELPOV)

-   Percent of housing units that are detached single family houses
    (PCTSINGLES)

-   Percent of housing units that are vacant (PCTVACANT)

-   Percent of residents in Block Group with at least a bachelor's
    degree (PCTBACHMOR)

Doing this gives us an cursory understanding of what the values our
predictors and dependent variable look like. Additionally we explored
the distribution of our data through the use of histograms. With a
histogram plot of the values of our variables, we can determine if they
are normally distributed. This is important for determining which kind
of regression we will use. Different regression models have different
assumptions. Some suggest that the residuals, or the estimate of error
of observations in our sample, for each variable should be normally
distributed.

It is possible for a non-normally distributed variable to have normally
distributed values, but it's much more likely that if the variable is
not normally distributed neither will its residuals. However, Central
Limit Theorem suggests that if you have enough observations, then the
normality of your variable or its residuals shouldn't matter. In any
case we are interested in exploring whether or not our variables are
distributed, so in the case that they are not, we can transform them and
preserve the assumptions of the regression we use. That is our intention
because we hope to make an interpretable model and having normally
distributed variables, and by extension residuals, our model will be
easier to interpret.

Correlation is a standardized measure that informs us how strong the
relationship is between two variables. This measure of relational
strength between variables is usually shown as *r*. The value of *r*
ranges from $-1 \leq r \geq 1$. A value of 1 or -1 implies a strong
correlation, but the slope and relationship will be negative with a
value of -1, and positive with a value of 1. A value of 0 implies there
is no linear relationship at all. It is possible that with a value of 0,
two variables do have some strong correlation, but it will not be linear
as the value *r* only calculates linear relationships.

For our data we used Pearson's Correlation equation:

$$r = \frac{{}\sum_{i=1}^{n} (x_i - \overline{x})(y_i - \overline{y})}
{\sqrt{\sum_{i=1}^{n} (x_i - \overline{x})^2(y_i - \overline{y})^2}}$$

Typically, if any two predictors have a correlation of $-0.8 < r > 0.8$
, then we might need to remove them from the model due to their
multicolinearity. Multicollinearity occurs when two variables are very
strongly correlated to each other. If two variables with strong
correlations were incorporated into the same model, there will be little
benefit and our coefficient estimates may be less reliable. Since we are
concerned with the interpretability of our model, we intended to avoid
multicollinearity.

###### **c. Multiple Regression Analysis**

We will use multiple Ordinary Least Squares (OLS) regression to
examine the relationship between our variable of interest and our
explanatory variables. OLS multiple regression determines the strength
and direction (positive, negative, zero) of the relationship, and
goodness of model fit. We calculate the coefficient 𝛽i of each predictor
which is interpreted as the amount by which the dependent variable, in
our case median house value, changes as the independent variable
increases by one unit, holding all other predictors constant. The model
includes the error term 𝜀, and is defined, for each observation i, as
the vertical distance from the observed value and the predicted value of
y. The error term is included in the equation to allow a point to fall
either above or below the regression line. We are regressing median
house value on percentage of homes living below poverty, percentage of
individuals with a bachelor's degree or higher, percentage of vacant
homes, and percentage of single house units in Philadelphia. Our
regression equation is as follows:

$$y = MEDHVAL = \beta_{0} + \beta_{1}LNBELPOV100 + \beta_{2}PCTBACHMOR + \beta_{3}PCTVACANT + \beta_{4}PCTSINGLES + \epsilon$$

OLS multiple regression models hold assumptions, the first being a
linear relationship between x and y, which was examined via our
scatter plots in Figure 3. The second assumption is normality of
residuals. This is important for point estimation, confidence intervals,
and hypothesis tests for small samples due to the Central Limit Theorem.
Normality is essential for all sample sizes in order to predict future
values of y. If residuals appear to have a non-normal distribution, it
may indicate a non-linear relationship, or non-normal distributions of
the dependent/independent variables. This may be solved with a
logarithmic transformation of the variables. The third and fourth
assumptions are that residuals are random and homoscedastic. There
should be no relationship or pattern between the observed values and the
predicted values. Homoscedasticity means the variance of residuals
should look constant for different values when plotted. The fifth
assumption states that observations and residuals must be independent.
If data has a temporal or spatial component, residuals of observations
that are close in time or space will be autocorrelated, or dependent on
one another. In this case, time series or spatial regression should be
used over OLS. The Final assumption of multiple OLS regression is the
absence of multicollinearity. Said differently, the predictors should
not be strongly correlated with one another. The presence of
multicollinearity causes unstable coefficient estimates and weakens the
model. The parameters of multiple regression are coefficients β~0~ ,...,
β~k~, where k = number of predictors. β~0~ represents the y- intercept of
the regression line. β~1~ ... β~k~ represent the coefficients of variables
x~1~ ... x~k~. Each independent variable will have its own slope coefficient
which will indicate the relationship of that predictor with the
dependent variable, controlling for all other independent variables in
the regression. OLS regression calculates the sum of the least squares of residuals.
The least squares estimates for β̂\_0,..., β ̂\_k are obtained when the quantity for 
SSE (equation below)is minimized. 

$$ 
SSE = \sum_{i=1}^{n} \epsilon^2 = \sum_{i=1}^{n} (y-\hat{y})^2    
= \sum_{i=1}^{n}(y_{i}- \hat{\beta}_{0}- \hat{\beta}_{1}x_{1i}- \hat{\beta}_{2}x_{2i}- ...- \hat{\beta}_{k}x_{ki})^2 
$$
Variance is the other parameter that needs to be estimated in multiple
regression, calculated as $σ^2=\frac{SSE}{(n-(k+1))}=MSE$ , where k =
number of predictors, n = number of observations, and MSE stands Mean
Squared Error. In multiple regression, R^2^ is the coefficient of multiple
determination, or the proportion of variance in the model explained by
all k predictors, represented by the equation $R^2=1-\frac{SSE}{SST}$ .
The R^2^ will increase with more predictors in the model and can be
adjusted for the number of predictors with the equation
$R_{adj}^2=\frac{(n-1) R^2-k}{n-(k+1)}$ The model utility test, F-ratio,
is conducted on the regression model to determine a goodness of fit
measure. It can be interpreted as a significance test for R^2^. The
F-ratio tests the null hypothesis, H~0~ that all coefficients in the model
are jointly zero, vs the alternative hypothesis Ha that at least one of
the coefficients is not 0. Said differently, we test to make sure none
of the independent variables is a significant predictor o the dependent
variable. From the F-ratio we are looking for a P value that is less
than 0.05. Once an F-ratio is determined, we run a T-test on every
individual predictor. The null hypothesis states that the predictor (ie;
Percentage of vacant homes) has no association with the dependent
variable, once again in our case is median house value. Our goal is to
reject the null hypothesis H~0~ in favor of the alternative hypothesis H~a~
which is Β~i~ ≠ 0 for each predictor.


###### **d. Additional Analysis**

#### **Results**

###### **a. Exploratory Results**

To gain a better understanding of our observations we calculated summary
statistics and examined the distribution of the variables. Below, Table
1 shows the summary statistics calculated for our dependent variable,
Median House Value (MEDHVAL) and our four independent variables:

-   Number of Households Living Below the Poverty Line (NBELPOV)

-   Percent of housing units that are detached single family houses
    (PCTSINGLES)

-   Percent of housing units that are vacant (PCTVACANT)

-   Percent of residents in Block Group with at least a bachelor's
    degree (PCTBACHMOR)

The distribution of our raw variables are shown in Figure 1. We thought
it wise to also illustrate the distribution of the natural log of all
our variables, as shown in Figure 2. From these visualizations, it was
clear that both the dependent variable and NBELPOV were not normally
distributed, and as a result were likely not great for our linear
regression. A linear regression model assumes that the residuals, or the
estimate of error of observations in our sample, for each variable is
normally distributed. And while it is possible for a variable with
non-normal distribution to have normally distributed residuals, it is
more likely that that variables with non-normal distributions also have
non-normally distributed residuals. If it were the case that the
residuals of our variables were not normally distributed, then we would
be in violation of the assumption of linear regression. Though there is
a case to be made, that we meet the criteria of Central Limit Theorem,
based on the amount of observations we have, we have decided to address
to non-normally distributed variables, so that we can confidently
perform statistically hypothesis testing with our model.

With this motivation, we concluded that it would be best to use the
natural logged dependent variable and NBELPOV. Additionally we chose to
use a natural log (and a natural log +1, where there were zeros in the
observations), because those two variables were positively skewed. We
used the raw versions of the remaining variables.

```{r summarydist_code, include=TRUE, results='hide', echo=FALSE, warning=FALSE, message=FALSE}

#### THIS FIRST CHUNK INCLUDES CODE FOR THE SUMMARY STATISTICS AND THE HISTOGRAMS ####

#save the summary to a df (i have 0 idea how this code works, just a copy paste from the web)
#we really only need the mean and sd, so I will clean this up
summary <-as.data.frame(apply(ourdata,2,summary))
#checking the index
base::row(summary,1)
#dropping all rows except the mean
summary<-summary[-c(1,2,3,5,6),]
#transposing so variables (mean and sd) are columns
summary <- summary%>%base::t()%>%as.data.frame()

#getting sd and adding sd column
sdMedV <- sd(ourdata$MEDHVAL)
sdBach <- sd(ourdata$PCTBACHMOR)
sdNbel <- sd(ourdata$NBELPOV100)
sdVac <- sd(ourdata$PCTVACANT)
sdSing <- sd(ourdata$PCTSINGLES)
sdMedHHINC <- sd(ourdata$MEDHHINC)

sdcol <- list(0, 0, sdMedV, sdBach, sdMedHHINC, sdVac, sdSing, sdNbel)
summary$sd <- sdcol

#dropping the other things we don't need 
base::row(summary,1)
summary<- summary[-c(1,2,5),]
summary<- summary%>%as.data.frame()
summary


#create a copy of the summary table to give it nicer row names for the table below
summary_nice_names <- summary
rownames(summary_nice_names) <- c("Median House Value",
                                  "% of Indviduals with Bachelor's Degrees or Higher",
                                  '% of Vacant Houses',
                                  "% of Single House Units",
                                  "% Households Living in Poverty")
#final summary stats table
summary_table <- kable(summary_nice_names) %>%
  kable_styling() %>%
  footnote(general_title = "Summary Statistics\n",
           general = "Table 1")


#---- Step 1A.ii: Histograms ----

#check them out, regular histogram

 
hists <- histogram( ~ MEDHVAL +PCTBACHMOR +NBELPOV100 +PCTSINGLES +PCTVACANT, layout=c(2,3), data = ourdata, main='Distribution of Raw Variables', sub= 'Figure 1', col="#8C96C6", breaks = 50, scales='free')   

#none look normal so we will examine if a log transformation will make them normal
#remember to use 1+ if any variable has zero values
#the only variable that does not have a 0 value is MEDHVAL

ourdata$LNMEDHVAL <- log(ourdata$MEDHVAL)

ourdata$LNPCTBACHMOR <- log(1+ourdata$PCTBACHMOR)

ourdata$LNBELPOV100 <- log(1+ourdata$NBELPOV100)

ourdata$LNPCTVACANT <- log(1+ourdata$PCTVACANT)

ourdata$LNPCTSINGLES <- log(1+ourdata$PCTSINGLES)


#### Logged 

LNhists <- histogram( ~ LNMEDHVAL +LNPCTBACHMOR +LNBELPOV100 +LNPCTSINGLES +LNPCTVACANT,layout=c(2,3),data = ourdata, 
                    main='Distribution of Natural Log of Variables', sub= 'Figure 2', col="#B3CDE3", breaks = 50, scales='free')  

```

```{r summary_output,  include = TRUE, echo= FALSE, warning=FALSE, message=FALSE}

summary_table
hists
LNhists


```

Next, we determined if there were any interesting relationships between
our predictors and our dependent variable. We were especially interested
in figuring out what kind of relationship our variables had, and whether
or not they were linear. Knowing this would help inform us what kind of
regression to use and what other measures we should use for exploring
our data. For example, as you'll see below we measured for correlation,
but we first needed to know if we could a measure for linear
relationships.

From our scatter plots in Figure 3, we can see that our predictors seems
to have a linear relationship with the natural log of our dependent
variable, Median House Value. These plots also give us an impression of
the relative strength of relationship between our predictors and the
dependent variable. But, our model will give us a clearer picture with
interpretable statistical meaning. For now, we are interested in
exploring the relationship between just our predictors, in order to rule
out any potential strong correlation between them.

```{r scatter_code, message=FALSE, echo=FALSE, results='hide', warning=TRUE, include=FALSE}

#### THIS CHUNK INCLUDE SCATTER PLOT CODE ####

#---- Step 1B: Variable Scatter Plots ----
#investigate to see if our predictors relationship with the dependent variable
# is linear by plotting as scatter plots

pLNBelpov <- ggplot(ourdata, aes(x = LNMEDHVAL, y= LNBELPOV100))+
            geom_point(size=2.5, color = "#4D004B", alpha = 0.5)+theme_minimal()
            
pBach <- ggplot(ourdata, aes(x = LNMEDHVAL, y= PCTBACHMOR))+
  geom_point(size=2.5, color = "#4D004B", alpha = 0.5)+theme_minimal()
  
  
pVac <- ggplot(ourdata, aes(x = LNMEDHVAL, y= PCTVACANT))+
  geom_point(size=2.5, color = "#4D004B", alpha = 0.5)+theme_minimal()
  
  

pSing <- ggplot(ourdata, aes(x = LNMEDHVAL, y= PCTSINGLES))+
  geom_point(size=2.5, color = "#4D004B", alpha = 0.5)+theme_minimal()+labs(caption="Figure 3")



scattergrid <- plot_grid( 
           pLNBelpov, 
           pBach, 
           pVac, 
           pSing, 
           labels = c("% Below Poverty (LN)", 
                      "Bachelors Degree",
                      "% Vacant Homes", 
                      "% Single House Units"),
           label_size = 11,
           label_x =-.1,
           label_y = 1.03,
           scale= 0.9,
           align = 'hv',
           ncol = 2, nrow = 2)



```

```{r scatter_output, include=TRUE, warning=FALSE, message=FALSE, echo=FALSE, fig.dim = c(10,12)}
scattergrid

```

Now knowing that we are interested in the linear relationships our
predictors have with our dependent variable, we need to investigate
whether there is a strong linear relationship between any of our
predictors. We investigated this using a correlation measure.

```{r corr_code_output, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

#### THIS CHUNK INCLUDE CORR CODE AND OUTPUT 
#---- Step 1C: Pearson Correlations ----
#make new df that just has our predictors and dependent variable

pred_var <- ourdata%>%dplyr::select(LNMEDHVAL, LNBELPOV100, PCTBACHMOR, PCTSINGLES, PCTVACANT)
pcorr <- cor(pred_var, method="pearson")

#Observe that there isnt severe multicollinearity (i.e., no correlations where
# r>.8 or r<-.8), so we can include all four predictors in the regression.

#*********matrix for markdown*****************************************
pcorr <- pcorr%>%kable(digits = c(4, 4, 4, 4, 4),
            align = c("l", "r", "r", "r", "r"))%>%kable_styling()%>%footnote(general_title = "Pearson's Correlation of Predictors",
             general = "Table 1")


pcorr
```

You'll notice that each predictor's correlation with itself is 1. But
none of our predictor variables have a correlation with another \> 0.8
or \<-0.8, so our model will be free from multicollinearity.

One last thing we looked at in our exploration of our data, is the
distribution of our predictor's values across space. Below are a series
of maps that show this distribution, with values broken out by Jenks
breaks. Map 1 shows us how our dependent variable, median house value,
differs across census tract, and Map 2 show our predictors variables
across tracts. Most have 5 breaks, but the values for LN %BELPOV were
note varied enough so we there are only three breaks.

```{r choropleth_code, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}

#### THIS CHUNK INCLUDES THE MAPPING & JENKS CODE ###

#---- Step 2A: Open Regression_Data shapefile (or GeoJSON in our case)----
#shapefiles are really hard to work with without having to download and mess with working directories in the code
#SO i converted the shapefile to a GeoJSON and will use that instead


ourdata_geom <- st_read("https://raw.githubusercontent.com/bri-ne/MUSA500_Assignment_1/main/RegressionData.geojson")


#getting Jenks Breaks for LNMEDHVAL 
classes <- classIntervals(ourdata_geom$LNMEDHVAL, n = 5, style = "jenks")
classes$brks

#we'll create a new column in our sf object using the base R cut() function to cut up our percent variable into distinct groups.

ourdata_geom <- ourdata_geom %>%
  mutate(LNMEDVHAL_class = cut(LNMEDHVAL, classes$brks, include.lowest = T))


#mapping
choro_LNMEDHVAL <- ggplot() +
  geom_sf(data = ourdata_geom,
          aes(fill = LNMEDVHAL_class),
          alpha = 1,
          colour = "gray80",
          size = 0.15) +
  scale_fill_brewer(palette = "PuBu",
                    name = "LN Median House Value") +
  labs(x = NULL, y = NULL,
       title = "LN Median House Value in Philadelphia by Block Group",
       subtitle = "Source: U.S. Census",
       caption = "Map 1") +
  theme(line = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank())

#making remaining JENKS for maps --------------------

#PctVacant Jenks
PVclasses <- classIntervals(ourdata_geom$PCTVACANT, n = 5, style = "jenks")

PVclasses$brks

ourdata_geom <- ourdata_geom %>%
  mutate(PCTVACANT_class = cut(PCTVACANT, PVclasses$brks, include.lowest = T))

#PctSingle Jenks
PSclasses <- classIntervals(ourdata_geom$PCTSINGLES, n = 5, style = "jenks")

PSclasses$brks

ourdata_geom <- ourdata_geom %>%
  mutate(PCTSINGLES_class = cut(PCTSINGLES, PSclasses$brks, include.lowest = T))

#PctBach Jenks
PBclasses <- classIntervals(ourdata_geom$PCTBACHMOR, n = 5, style = "jenks")

PBclasses$brks

ourdata_geom <- ourdata_geom %>%
  mutate(PCTBACHMOR_class = cut(PCTBACHMOR, PBclasses$brks, include.lowest = T))

#LNBelPov Jenks
BPclasses <- classIntervals(ourdata_geom$LNNBELPOV, n = 3, style = "jenks")

BPclasses$brks
ourdata_geom <- ourdata_geom %>%
  mutate(LNNBELPOV_class = cut(LNNBELPOV, BPclasses$brks, include.lowest = T))


#mapping the rest --------------------------------------

#PctVacant Map

choro_PctVac <- ggplot() +
  geom_sf(data = ourdata_geom,
          aes(fill = PCTVACANT_class),
          #alpha = 1,
          colour = NA) +
  scale_fill_brewer(palette = "PuBu",
                    name = "Percent Houses Vacant") +
  labs(x = NULL, y = NULL,
       subtitle = "Perecent of Vacant Houses\nin Philadelphia by Block Group") +
  theme(line = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank())



#PctSing Map
choro_PctSing <- ggplot() +
  geom_sf(data = ourdata_geom,
          aes(fill = PCTSINGLES_class),
          #alpha = 1,
          colour = NA) +
  scale_fill_brewer(palette = "PuBu",
                    name = "Percent Single House Units") +
  labs(x = NULL, y = NULL,
       subtitle = "Perecent of Single House Units\nin Philadelphia by Block Group") +
  theme(line = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank())


#PctBach Map
choro_PctBach <- ggplot() +
  geom_sf(data = ourdata_geom,
          aes(fill = PCTBACHMOR_class),
          #alpha = 1,
          colour = NA) +
  scale_fill_brewer(palette = "PuBu",
                    name = "Percent Bachelors Degree") +
  labs(x = NULL, y = NULL,
       subtitle = "Percent of Individuals with a Bachelors\nDegree or Higher in Philadelphia\nby Block Group") +
  theme(line = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank())

#LNBelPov Map
choro_LNBelPov <- ggplot() +
  geom_sf(data = ourdata_geom,
          aes(fill = LNNBELPOV_class),
          #alpha = 1,
          colour = NA) +
  scale_fill_brewer(palette = "PuBu",
                    name = "LN Percent Households\nin Poverty") +
  labs(x = NULL, y = NULL,
       subtitle = "LN Percent of Households Living\nin Poverty in Philadelphia\nby Block Group",
       caption = "Map 2. Source: U.S. Census") +
  theme(line = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank())

#--- Unfortunately I had to do three breaks b/c the values for the LN %BELPOV were not varied enough


mapgrid <- plot_grid( choro_PctVac, 
                      choro_PctSing, 
                      choro_PctBach, 
                      choro_LNBelPov,
                      align = c("hv","hv","hv","hv"),
                          ncol = 2, nrow = 2)

```

```{r maps_output, include=TRUE, echo=FALSE, fig.dim=c(10,12), warning=FALSE, message=FALSE}

choro_LNMEDHVAL
mapgrid
```

The map of the percent individuals with a Bachelors or higher looks the
most similar to our dependent variable, while the percent of households
living in poverty looks almost the opposite. It also appears as though
where there are people with at least Bachelors, there are also single
house units and house value is higher.

###### **b. Regression Results**

After we explore our data and transform our variables as needed, the final regression equation is as follows:

$$y = LNMEDHVAL = \beta_{0} + \beta_{1}LNBELPOV100 + \beta_{2}PCTBACHMOR + \beta_{3}PCTVACANT + \beta_{4}PCTSINGLES + \epsilon$$
We regressed the natural log of median house value (LNMEDHVAL) on the natural log of the number of people living below poverty (LNBELPOV100), the percent of individuals with Bachelor's Degrees or higher (PCTBACHMORE), the percent of vacant houses (PCTVACANT), and the percent of single house units (PCTSINGLES). The regression output tells  us  that  these variables are highly significant predictors and are positively associated with the median house value (p<0.0001 for all variables). The coefficients will have a modified interpretation because we have used a logarithmic transformation on the dependent variable. To interpret the beta coefficient of LNBELPOV100, we can say a 1% increase in the number of individuals living in poverty corresponds to a $(1.01^{\beta_1} - 1)*100 = (1.01^{-.079} - 1) * 100 = -0.0786 \%$ change (i.e., a .0786% decrease) in median house value. The regression also tells us that a 1% increase in the percent of individuals with a Bachelor's degree or more is associated with a change in median house value by $\beta_2 ln(\frac {101}{100}) = 0.0209 ln (\frac {101}{100}) = 0.00021$ \$ (i.e., an increase in median house value by 0.00021 \$). The  next coefficients will be interpreted in a similar way. A 1% increase in percent of vacant homes corresponds to an approximately $\beta_3 ln(\frac {101}{100}) = 0.0191 ln (\frac {101}{100}) = 0.00019$ \$ increase in median house value. Lastly, a 1% increase in percent of single unit homes corresponds to an approximately $\beta_4 ln(\frac {101}{100}) = 0.0029 ln (\frac {101}{100}) = 0.00003$ \$ increase in median house value. 

The p value of less than 0.0001 for LNBELPOV100 tells us that if there is actually no relationship between LNBELPOV100 and the dependent variable LNMEDHVAL (i.e., if  the  null  hypothesis  that  $\beta_1 =0$  is  actually  true),  then  the  probability  of  getting  a $\beta_1$ coefficient  estimate  of  -0.0789054  is  less  than  0.0001.  Similarly,  the  p-value  of  less  than 0.0001 for PCTBACHMORE tells us that *if* there is actually no relationship between PCTBACHMORE and the dependent variable MEDHVAL (i.e., if the null hypothesis that $\beta_2=0$ is actually true), then the probability of getting a $\beta_2$ coefficient estimate of 0.02091 is less than 0.0001. The same interpretations can be made for the p values of PCTVACANT and PCTSINGLES - both of these predictors are statistically significant with a very low p value < 0.0001. These low probabilities indicate that we can safely reject  
$H_0: \beta_1 = 0$ for $H_a: \beta_1 ≠ 0$,   
$H_0: \beta_2 = 0$ for $H_a: \beta_2 ≠ 0$,   
$H_0: \beta_3 = 0$ for $H_a: \beta_3 ≠ 0$,   
$H_0: \beta_4 = 0$ for $H_a: \beta_4 ≠ 0$   
(at most reasonable levels of α = P(Type I error)).

A over half of the variance in the dependent variable is explained by the model (R2 and Adjusted R2 are 0.6623 and 0.6615, respectively). The low p-value associated with the F-ratio shows that we can reject the null hypothesis that all coefficients in the model are 0.

```{r regressionanalysis, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE, results='hide'}

#---- Step 3 : Regression Analysis -------

#Run regression model
#fit <- lm(MEDHVAL ~ LNBELPOV100 + PCTBACHMOR + PCTVACANT + PCTSINGLES, data=ourdata)
fit <- lm(LNMEDHVAL ~ LNBELPOV100 + PCTBACHMOR + PCTVACANT + PCTSINGLES, data=ourdata)

summary(fit)
#3A

fit.summary<- broom::tidy(fit)%>%
              kable(
              align = c("l", "r", "r", "r", "r"))%>%kable_styling()%>%footnote(general_title = "Regression Summary",
              general = "Table 2")

#displays the summary statistics and R squared
fit.stats<- glance(fit)%>%kable()%>%kable_styling()%>%footnote(general_title = "Regression Summary",
              general = "Table 3")

#Summary of Regression
#all predictors are significant - p value < 0.05
#R squared - 55.38% = % of variance in median house value explained 
    #by predictors 
#Adjusted R Squared - 55.28% = % of variance in median house value explained
    #BY predictors adjusted for the number of predictors
#P value associated with F-ratio of 532 is less than 0.0001 -
    #We can reject the H0 that all Beta coefficients for the predictors are 0

#3B
anova.fit<- broom::tidy(anova(fit))%>%kable(digits = c(0, 4, 4, 4, 4),
            align = c("l", "r", "r", "r", "r", "r"))%>%kable_styling()%>%footnote(general_title = "ANOVA",
             general = "Table 4")

#ANOVA table containing SSE and SSR
#SSR = SS(LNBELPOV100) + SS(PCTBACHMOR) +SS(PCTVACANTS) + SS(PCTSINGLES) = 
    # 869866100108 + 2345334804156 +  58030777316 + 154801231763 = 3428032913343
    # Amount of total variance in Median House value explained by model
#SSE = 2761620505240 = amount of total variance in median house value that is 
    # unexplained by the model

#3C
#new column saving predicted values 
ourdata$predvals <- fitted(fit)

#new column saving residuals
ourdata$resids <- residuals(fit)

#New column saving standardized residuals
ourdata$stdres <- rstandard(fit)

#3D Create a scatterplot with standardized residuals on Y axis and
    # Predicted Values on x axis. 

ResidualPlot<- ggplot(ourdata, aes(x = predvals, y= stdres))+
                  geom_point(size=1.5, color = 'darkslateblue', alpha = .5)+
                  geom_hline(yintercept=0, size = .5)+
                  labs(x = 'Predicted LN Median House Value', 
                       y = 'Standardized Residuals',
                       caption = 'Figure 4')

```


```{r summary.anova, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}
fit.summary
fit.stats
anova.fit

```

```{r residplot, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
ResidualPlot
```


###### **c. Regression Assumption Checks**

###### **d. Additional Models**

#### **Limitations**
